---
title: "Problem 3: Topic Modeling with LDA in R"
author: "Stephen Pasco"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    highlight: tango
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6
)
```

# Overview

This notebook implements LDA topic modeling using R, replicating the analysis from Problem 1 but using R's `topicmodels` package instead of Python's Gensim.

**Objective:** Extract topics from an academic article about network-knowledge graph duality in supply chain risk analysis.

---

# (a) Data Loading and Preparation

## Install and Load Required Packages

```{r load-packages}
# Install required packages if not already installed
required_packages <- c("tm", "topicmodels", "SnowballC", "jsonlite", "knitr")

for (pkg in required_packages) {
  if (!require(pkg, character.only = TRUE, quietly = TRUE)) {
    cat(sprintf("Installing package: %s\n", pkg))
    install.packages(pkg, repos = "https://cloud.r-project.org/", quiet = TRUE)
    library(pkg, character.only = TRUE)
  }
}
```

## Load Article Data

```{r load-data}
# Load article data from JSON
article_data <- fromJSON("article-sections.json")

cat(sprintf("Article: 'Network-Knowledge Graph Duality in Supply Chain Risk Analysis'\n"))
cat(sprintf("Total sections (documents): %d\n", nrow(article_data)))
cat(sprintf("Number of topics to extract: 6 (same as Problem 1)\n"))
```

## Extract Documents

```{r extract-documents}
# Extract document content
documents <- article_data$content
section_headers <- article_data$section

# Show sample of first few sections
kable(
  data.frame(
    Section = section_headers[1:5],
    Preview = substr(documents[1:5], 1, 100)
  ),
  caption = "Sample of First 5 Document Sections"
)
```

---

# (b) Text Preprocessing

```{r preprocessing}
library(tm)
library(SnowballC)

cat("Preprocessing steps:\n")
cat("  - Convert to lowercase\n")
cat("  - Remove numbers and punctuation\n")
cat("  - Remove English stopwords\n")
cat("  - Apply stemming\n")
cat("  - Filter words with length < 3\n\n")

# Create corpus
corpus <- Corpus(VectorSource(documents))

# Create document-term matrix with preprocessing controls
dtm <- DocumentTermMatrix(corpus, control = list(
  tolower = TRUE,
  removeNumbers = TRUE,
  removePunctuation = TRUE,
  stopwords = stopwords("english"),
  stemming = TRUE,
  wordLengths = c(3, Inf)
))

cat(sprintf("Initial DTM: %d documents x %d terms\n", nrow(dtm), ncol(dtm)))
```

## Filter Sparse Terms

```{r filter-sparse}
# Remove sparse terms (similar to Problem 1's filtering)
dtm_filtered <- removeSparseTerms(dtm, 0.7)

# Remove empty documents (rows with all zeros)
row_sums <- rowSums(as.matrix(dtm_filtered))
dtm_filtered <- dtm_filtered[row_sums > 0, ]
documents_filtered <- documents[row_sums > 0]
section_headers_filtered <- section_headers[row_sums > 0]

cat(sprintf("Filtered DTM: %d documents x %d terms\n", nrow(dtm_filtered), ncol(dtm_filtered)))
cat(sprintf("Documents retained after filtering: %d\n", nrow(dtm_filtered)))
```

---

# (c) LDA Model Training

```{r train-lda}
library(topicmodels)
set.seed(42)  # For reproducibility

cat("Training LDA model with 6 topics...\n")
lda_model <- LDA(
  dtm_filtered,
  k = 6,
  method = "VEM",
  control = list(
    seed = 42,
    estimate.alpha = TRUE,
    verbose = 0
  )
)
cat("Model trained successfully.\n")
```

---

# (d) Results Presentation

```{r helper-function}
# Helper function to generate topic labels
get_topic_label <- function(top_words) {
  word_str <- paste(tolower(top_words[1:7]), collapse = " ")

  if (grepl("agent|triage|tool|call|user|queri", word_str)) {
    return("Agent Systems")
  } else if (grepl("node|edg|path|central|travers", word_str)) {
    return("Network Paths")
  } else if (grepl("factor|portfolio|secur|stock|shell", word_str)) {
    return("Financial Analysis")
  } else if (grepl("suppli|chain|coltan|appl|drc", word_str)) {
    return("Supply Chain")
  } else if (grepl("model|time|domain|hop|summari", word_str)) {
    return("Graph Traversal")
  } else if (grepl("graph|network|llm|retriev", word_str)) {
    if (grepl("llm|languag", word_str)) {
      return("LLM Applications")
    } else {
      return("Knowledge Graphs")
    }
  } else {
    return(paste(tools::toTitleCase(top_words[1:2]), collapse = " "))
  }
}
```

```{r extract-results}
# Get top terms and document-topic distributions
top_terms <- terms(lda_model, 10)
doc_topics <- posterior(lda_model)$topics
topic_term_probs <- posterior(lda_model)$terms
```

## Topic Analysis

```{r topics-display, results='asis'}
# Display results for each topic
for (topic_id in 1:6) {
  cat(sprintf("\n## Topic %d\n\n", topic_id - 1))  # 0-indexed to match Problem 1

  # Top 10 words
  cat("### Top 10 Most Frequent Words\n\n")
  topic_words <- top_terms[, topic_id]

  word_table <- data.frame(
    Rank = 1:10,
    Word = character(10),
    Probability = numeric(10),
    stringsAsFactors = FALSE
  )

  for (rank in 1:10) {
    word <- topic_words[rank]
    word_idx <- which(colnames(topic_term_probs) == word)
    if (length(word_idx) > 0) {
      prob <- topic_term_probs[topic_id, word_idx]
      word_table$Word[rank] <- word
      word_table$Probability[rank] <- prob
    }
  }

  print(kable(word_table, digits = 4))

  # Topic label
  label <- get_topic_label(topic_words)
  cat(sprintf("\n**Topic Label:** %s\n\n", label))

  # Top 2 most associated documents
  topic_scores <- doc_topics[, topic_id]
  top_doc_indices <- order(topic_scores, decreasing = TRUE)[1:2]

  cat("### Top 2 Most Associated Documents\n\n")
  for (rank in 1:2) {
    doc_idx <- top_doc_indices[rank]
    score <- topic_scores[doc_idx]
    header <- section_headers_filtered[doc_idx]
    content <- documents_filtered[doc_idx]
    preview <- if (nchar(content) > 150) {
      paste0(substr(content, 1, 150), "...")
    } else {
      content
    }

    cat(sprintf("**Document %d** (Association Score: %.4f)\n\n", rank, score))
    cat(sprintf("- **Section:** %s\n", header))
    cat(sprintf("- **Preview:** %s\n\n", preview))
  }

  cat("\n---\n\n")
}
```

---

# Conclusion

This R implementation of LDA topic modeling successfully extracted 6 topics from the academic article about network-knowledge graph duality in supply chain risk analysis.

**Note:** Topics derived in R may differ from those in Problem 1 due to:

- Differences in library implementations (R's `topicmodels` vs Python's Gensim)
- Preprocessing differences (stemming vs lemmatization)
- Random initialization and optimization algorithms

Both implementations follow the same conceptual approach but may produce different topic distributions due to these technical variations.

---

# Session Information

```{r session-info}
sessionInfo()
```
