---
title: "Assignment 7 - Structural Topic Modeling (Problems 1 & 2)"
author: "Stephen Pasco"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 6)
```

```{r load-packages}
library(stm)
library(SnowballC)
library(wordcloud)
library(tm)
library(ggplot2)
library(scales)
```

```{r helper-functions}
# Helper function: Min-max normalization to [0, 1] scale
# This function rescales a numeric vector to the range [0, 1]
# Used for combining metrics with different scales (e.g., coherence and exclusivity)
normalize_metric <- function(x) {
  x_min <- min(x, na.rm = TRUE)
  x_max <- max(x, na.rm = TRUE)

  # Handle edge case where all values are identical
  if (x_max == x_min) {
    return(rep(0.5, length(x)))
  }

  (x - x_min) / (x_max - x_min)
}
```

---

# Part A: Data Preparation

## Loading the Dataset

```{r load-data}
# Load the hotel reviews dataset
hotel_data <- read.csv("data/hotel_reviews.csv", stringsAsFactors = FALSE)

# Validate data integrity
stopifnot(
  "Data file must contain Review column" = "Review" %in% names(hotel_data),
  "Data file must contain Rating column" = "Rating" %in% names(hotel_data),
  "Dataset must not be empty" = nrow(hotel_data) > 0,
  "Reviews must not be empty strings" = all(nchar(trimws(hotel_data$Review)) > 0),
  "Ratings must be between 1 and 5" = all(hotel_data$Rating >= 1 & hotel_data$Rating <= 5)
)

cat("Dataset loaded successfully:", nrow(hotel_data), "reviews\n")
```

## Text Preprocessing

```{r preprocess}
# Extract reviews and ratings
reviews <- hotel_data$Review
ratings <- hotel_data$Rating

# Preprocess text using textProcessor
processed <- textProcessor(
  documents = reviews,
  metadata = data.frame(rating = ratings),
  lowercase = TRUE,
  removestopwords = TRUE,
  removenumbers = TRUE,
  removepunctuation = TRUE,
  stem = TRUE,
  wordLengths = c(3, Inf)
)

# Prepare documents
# lower.thresh = 10 removes terms appearing in < 10 documents (~0.05% of corpus)
# This filters out extremely rare terms that may be typos or noise while
# preserving domain-specific vocabulary
out <- prepDocuments(
  documents = processed$documents,
  vocab = processed$vocab,
  meta = processed$meta,
  lower.thresh = 10
)

docs <- out$documents
vocab <- out$vocab
meta <- out$meta
```

---

# Part B: STM Model Implementation

## Running selectModel()

We run `selectModel()` to fit multiple STM models with 3 topics, using Rating as a prevalence covariate. The function performs multiple random initializations to help us select the best model based on semantic coherence and exclusivity.

```{r select-model, results='hide'}
set.seed(42)

# Run selectModel with 3 topics and Rating as covariate
models <- selectModel(
  documents = docs,
  vocab = vocab,
  K = 3,
  prevalence = ~ rating,
  max.em.its = 75,
  data = meta,
  runs = 10,
  seed = 42
)
```

## Model Selection: Semantic Coherence vs Exclusivity

Semantic coherence measures how often the top words in a topic co-occur in documents (higher is better). Exclusivity measures how distinctive a topic's words are compared to other topics (higher is better). We want a model that balances both metrics.

```{r plot-model-metrics}
# Plot semantic coherence vs exclusivity
plotModels(models)
title("Model Selection: Semantic Coherence vs Exclusivity")
```

## Select the Best Model

We select the model with the best balance between semantic coherence and exclusivity. Typically, we want high values for both, but there's often a trade-off.

```{r select-best-model}
# Select the model with highest average of normalized coherence and exclusivity
coherence <- unlist(lapply(models$runout, function(x) mean(semanticCoherence(x, docs))))
exclusivity <- unlist(lapply(models$runout, function(x) mean(exclusivity(x))))

# Normalize to [0,1] scale using helper function
coherence_norm <- normalize_metric(coherence)
exclusivity_norm <- normalize_metric(exclusivity)

# Combined score (equal weighting)
combined_score <- coherence_norm + exclusivity_norm
best_model_idx <- which.max(combined_score)

# Extract the best model
stm_model <- models$runout[[best_model_idx]]
```

---

# Part C: Results Analysis

## Top 7 Words for Each Topic

```{r top-words}
# Extract top words for each topic
top_words <- labelTopics(stm_model, n = 7)

cat("Top 7 words for each topic:\n\n")
for (i in 1:3) {
  cat("Topic", i, ":\n")
  cat("  FREX words:", paste(top_words$frex[i,], collapse = ", "), "\n")
  cat("  Highest prob:", paste(top_words$prob[i,], collapse = ", "), "\n\n")
}
```

## Word Clouds for Each Topic

Word clouds visualize the most prevalent words in each topic, with size reflecting word importance.

```{r wordclouds, fig.height=4, fig.width=12}
# Create word clouds for each topic
par(mfrow = c(1, 3))
for (i in 1:3) {
  cloud(stm_model, topic = i, scale = c(3, 0.5), max.words = 50)
  title(paste("Topic", i), line = 1)
}
par(mfrow = c(1, 1))
```

## Top 5 Reviews for Each Topic

These are the reviews most strongly associated with each topic.

```{r top-reviews}
# Find top reviews for each topic
for (i in 1:3) {
  cat("\n", rep("=", 80), "\n", sep = "")
  cat("Topic", i, "- Top 5 Reviews:\n")
  cat(rep("=", 80), "\n\n", sep = "")

  thoughts <- findThoughts(stm_model, texts = reviews, topics = i, n = 5)

  # Get the indices (thoughts$index is a list)
  indices <- thoughts$index[[1]]

  # Safety check: ensure we have valid data
  n_available <- min(5, length(indices), length(thoughts$docs[[1]]))

  if (n_available == 0) {
    cat("No reviews found for this topic.\n\n")
    next
  }

  for (j in 1:n_available) {
    cat("Review", j, "(Rating:", meta$rating[indices[j]], "):\n")
    # Show first 400 characters of each review (word-boundary aware)
    review_text <- as.character(thoughts$docs[[1]][j])
    truncated <- substr(review_text, 1, 400)
    # Add ellipsis only if text was actually truncated
    if (nchar(review_text) > 400) {
      cat(truncated, "...\n\n")
    } else {
      cat(review_text, "\n\n")
    }
  }
}
```

## Topic Labels

Based on the top words and representative reviews above, assign interpretive labels:

```{r topic-labels}
topic_labels <- c(
  "Topic 1" = "Room Quality and Service Complaints",
  "Topic 2" = "Beach Resort Amenities and Activities",
  "Topic 3" = "City Hotel Location and Access"
)

for (i in 1:3) {
  cat("Topic", i, ":", topic_labels[i], "\n")
}
```

---

# Part D: Visualization and Interpretation

## Topic Prevalence

This plot shows the overall prevalence (frequency) of each topic across all reviews.

```{r topic-prevalence}
# Plot topic prevalence
plot(stm_model, type = "summary", text.cex = 0.8, main = "Topic Prevalence in Hotel Reviews")
```

## Effect of Rating on Topic Prevalence

We use `estimateEffect()` to model how ratings influence the prevalence of each topic.

```{r estimate-effects, results='hide'}
# Estimate the effect of rating on topic prevalence
effects <- estimateEffect(
  formula = 1:3 ~ rating,
  stmobj = stm_model,
  metadata = meta,
  uncertainty = "Global"
)
```

## Visualizing Rating Effects

These plots show how topic prevalence changes with different ratings.

```{r plot-effects, fig.height=4, fig.width=12}
# Plot the effect of rating on each topic
par(mfrow = c(1, 3))
for (i in 1:3) {
  plot(
    effects,
    covariate = "rating",
    topics = i,
    model = stm_model,
    method = "continuous",
    xlab = "Rating (1-5)",
    main = paste("Effect of Rating on", topic_labels[i]),
    xlim = c(1, 5),
    printlegend = FALSE
  )
}
par(mfrow = c(1, 1))
```

## What Customers Value vs Dislike

```{r interpretation}
for (topic_num in 1:3) {
  topic_summary <- summary(effects, topics = topic_num)
  rating_coef <- topic_summary$tables[[1]][2, "Estimate"]
  rating_pval <- topic_summary$tables[[1]][2, "Pr(>|t|)"]

  cat("Topic", topic_num, "- Rating Coefficient:", round(rating_coef, 5),
      "(p =", format.pval(rating_pval, digits = 3), ")")

  if (rating_coef > 0) {
    cat(" - Valued by customers\n")
  } else {
    cat(" - Disliked by customers\n")
  }
}
```

---

# Part E: Problem 2 - Determining Optimal Number of Topics with searchK()

## Overview

In Problem 1, we used `selectModel()` with a fixed K=3 topics. Now in Problem 2, we systematically determine the optimal number of topics using `searchK()`, which evaluates models across different values of K by computing diagnostic metrics.

## (a) Running searchK() for K = 2 to 6

```{r search-k, results='hide'}
set.seed(42)

# Search for optimal K from 2 to 6 topics
search_results <- searchK(
  documents = docs,
  vocab = vocab,
  K = c(2, 3, 4, 5, 6),
  prevalence = ~ rating,
  data = meta,
  max.em.its = 75,
  init.type = "Spectral",
  seed = 42
)

# Validate searchK() results
cat("searchK() completed for", nrow(search_results$results), "K values\n")
cat("Expected:", length(c(2, 3, 4, 5, 6)), "K values\n")
```

## (b) Exclusivity vs Semantic Coherence

Semantic coherence measures how often the top words in a topic co-occur in documents (higher is better). Exclusivity measures how distinctive a topic's words are compared to other topics (higher is better).

```{r plot-search-results, fig.height=6}
# Use the built-in plot function from STM package
plot(search_results)
```

---

# Part F: Composite Score Analysis

## (c) Rescaling Metrics to [0, 1]

We normalize both coherence and exclusivity to the [0, 1] scale:

```{r rescale-metrics}
# Extract metrics from search_results using unlist()
metrics_df <- data.frame(
  K = c(2, 3, 4, 5, 6),
  Coherence = unlist(search_results$results$semcoh),
  Exclusivity = unlist(search_results$results$exclus)
)

# Normalize coherence and exclusivity to [0, 1] scale using helper function
metrics_df$Coherence_Norm <- normalize_metric(metrics_df$Coherence)
metrics_df$Exclusivity_Norm <- normalize_metric(metrics_df$Exclusivity)

# Display the rescaled metrics
cat("Rescaled Metrics:\n")
print(metrics_df[, c("K", "Coherence_Norm", "Exclusivity_Norm")])
```

## (d) Calculate and Plot Composite Score

**Composite Score = 0.5 × (Rescaled Coherence) + 0.5 × (Rescaled Exclusivity)**

```{r calculate-composite}
# Calculate composite score with equal weighting (0.5 each)
metrics_df$Composite_Score <- 0.5 * metrics_df$Coherence_Norm + 0.5 * metrics_df$Exclusivity_Norm

# Display the complete metrics table
cat("Metrics Summary with Composite Score:\n")
print(metrics_df)
```

```{r plot-composite-components, fig.height=6, fig.width=10}
# Plot showing all three metrics: rescaled coherence, rescaled exclusivity, and composite score
ggplot(metrics_df, aes(x = K)) +
  geom_line(aes(y = Coherence_Norm, color = "Rescaled Coherence"), linewidth = 1.2) +
  geom_point(aes(y = Coherence_Norm, color = "Rescaled Coherence"), size = 4) +
  geom_line(aes(y = Exclusivity_Norm, color = "Rescaled Exclusivity"), linewidth = 1.2) +
  geom_point(aes(y = Exclusivity_Norm, color = "Rescaled Exclusivity"), size = 4) +
  geom_line(aes(y = Composite_Score, color = "Composite Score"), linewidth = 1.2) +
  geom_point(aes(y = Composite_Score, color = "Composite Score"), size = 4) +
  scale_color_manual(
    values = c(
      "Rescaled Coherence" = "steelblue",
      "Rescaled Exclusivity" = "coral",
      "Composite Score" = "darkgreen"
    )
  ) +
  scale_x_continuous(breaks = metrics_df$K) +
  labs(
    title = "Composite Score and Its Components vs Number of Topics",
    subtitle = "Composite Score = 0.5 × (Rescaled Coherence) + 0.5 × (Rescaled Exclusivity)",
    x = "Number of Topics (K)",
    y = "Normalized Score [0, 1]",
    color = "Metric"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 11),
    axis.title = element_text(size = 12),
    legend.position = "bottom"
  )
```

---

## (e) Plot Composite Score and Select Optimal K

```{r plot-composite-score, fig.height=5}
# Create composite score plot focused on optimal K selection
ggplot(metrics_df, aes(x = K, y = Composite_Score)) +
  geom_line(color = "darkblue", linewidth = 1.2) +
  geom_point(size = 4, color = "steelblue") +
  geom_point(
    data = metrics_df[which.max(metrics_df$Composite_Score), ],
    aes(x = K, y = Composite_Score),
    size = 6,
    color = "red",
    shape = 1,
    stroke = 2
  ) +
  scale_x_continuous(breaks = metrics_df$K) +
  labs(
    title = "Composite Score vs Number of Topics",
    subtitle = "Optimal K highlighted in red",
    x = "Number of Topics (K)",
    y = "Composite Score"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 11),
    axis.title = element_text(size = 12)
  )
```

### Justification for Optimal K

```{r select-optimal-k}
optimal_k <- metrics_df$K[which.max(metrics_df$Composite_Score)]
optimal_score <- max(metrics_df$Composite_Score)

cat(
  "\n**OPTIMAL NUMBER OF TOPICS:** K =", optimal_k, "\n\n",
  "**Composite Score:**", round(optimal_score, 4), "\n\n",
  "**JUSTIFICATION:**\n\n",
  "K =", optimal_k, "maximizes the composite score (", round(optimal_score, 4),
  "), which equally weights rescaled semantic coherence (how often top words co-occur in documents) and exclusivity (how distinctive topics are from each other). This represents the optimal trade-off between interpretable topics and minimal topic overlap.\n\n"
)
```

---

# Part G: Final Model Results with Optimal K

## (f) Results for Optimal K

### Fitting STM Model with Optimal K

```{r fit-optimal-model, results='hide'}
# Fit the final model with optimal K
optimal_model <- stm(
  documents = docs,
  vocab = vocab,
  K = optimal_k,
  prevalence = ~ rating,
  data = meta,
  max.em.its = 75,
  seed = 42,
  init.type = "Spectral"
)
```

### Top 7 Words for Each Topic

```{r optimal-top-words}
# Extract top words for each topic
optimal_top_words <- labelTopics(optimal_model, n = 7)

cat("Top 7 words for each topic (K =", optimal_k, "):\n\n")
for (i in 1:optimal_k) {
  cat("Topic", i, ":\n")
  cat("  FREX words:    ", paste(optimal_top_words$frex[i,], collapse = ", "), "\n")
  cat("  Highest prob:  ", paste(optimal_top_words$prob[i,], collapse = ", "), "\n\n")
}
```

### Word Clouds for Each Topic

```{r optimal-wordclouds, fig.height=4, fig.width=12}
# Determine layout based on number of topics
if (optimal_k <= 3) {
  par(mfrow = c(1, optimal_k))
} else if (optimal_k == 4) {
  par(mfrow = c(2, 2))
} else if (optimal_k <= 6) {
  par(mfrow = c(2, 3))
}

# Create word clouds for each topic
for (i in 1:optimal_k) {
  cloud(optimal_model, topic = i, scale = c(3, 0.5), max.words = 50)
  title(paste("Topic", i), line = 1)
}

par(mfrow = c(1, 1))
```

### Top Review for Each Topic

```{r optimal-top-reviews}
# Find top review for each topic
for (i in 1:optimal_k) {
  cat("\n", rep("=", 80), "\n", sep = "")
  cat("Topic", i, "- Top Review:\n")
  cat(rep("=", 80), "\n\n", sep = "")

  thoughts <- findThoughts(optimal_model, texts = reviews, topics = i, n = 1)

  # Safety check: ensure we have valid data
  if (length(thoughts$index[[1]]) == 0 || length(thoughts$docs[[1]]) == 0) {
    cat("No reviews found for this topic.\n\n")
    next
  }

  # Get the index
  index <- thoughts$index[[1]][1]

  cat("Rating:", meta$rating[index], "\n\n")
  cat("Review:\n")
  cat(as.character(thoughts$docs[[1]][1]), "\n\n")
}
```

### Topic Prevalence for Optimal Model

```{r optimal-prevalence}
# Plot topic prevalence
plot(optimal_model, type = "summary", text.cex = 0.8,
     main = paste("Topic Prevalence in Hotel Reviews (K =", optimal_k, ")"))
```

---

# Part H: Summary and Comparison

## Comparison: Problem 1 (K=3) vs Problem 2 (Optimal K)

```{r comparison}
conclusion <- if (optimal_k == 3) {
  "**CONCLUSION:** The data-driven approach confirms that K = 3 is optimal.\n"
} else {
  paste0("**CONCLUSION:** The data-driven approach suggests K = ", optimal_k, " provides better balance than the fixed K = 3 used in Problem 1.\n")
}

cat(
  "\n**PROBLEM 1: Fixed K = 3 topics**\n\n",
  "- Used selectModel() with multiple random initializations\n",
  "- Selected best model based on semantic coherence and exclusivity\n\n",
  "---\n\n",
  "**PROBLEM 2: Data-driven optimal K =", optimal_k, "topics**\n\n",
  "- Used searchK() to evaluate K = 2, 3, 4, 5, 6\n",
  "- Selected K based on composite score balancing coherence and exclusivity\n",
  "- **Composite Score:**", round(optimal_score, 4), "\n\n",
  "---\n\n",
  conclusion
)
```

---

# Reproducibility Information

## Session Information

This section documents the R environment, package versions, and system information to ensure reproducibility of results.

```{r session-info}
# Display complete session information including R version and loaded packages
sessionInfo()
```

## Data Summary

```{r data-summary}
cat(
  "**Dataset:** hotel_reviews.csv\n",
  "**Total Reviews:**", nrow(hotel_data), "\n",
  "**Rating Distribution:**\n"
)
table(hotel_data$Rating)

cat("\n**Processing Summary:**\n")
cat("- Original vocabulary size:", length(processed$vocab), "unique terms\n")
cat("- After filtering (lower.thresh = 10):", length(vocab), "unique terms\n")
cat("- Documents retained:", length(docs), "out of", length(reviews), "\n")
```
